---
title: "Brownlows Bets"
author: "Racheal Adeneye"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->
###First, we will run the brlw function 1000 seperate times and each time it runs, we will capture the number that represents how much Brownlow will make. Next we will sort the 1000 different numbers from low to high. Then, we will grab the lower two and a half percent.

```{r}
MyData= numeric(1000)
for (i in 1:1000) {
  MyData[i]=BrLw()
}
sMyData =sort(MyData)
low=sMyData[25]
high=sMyData[975]
average=mean(sMyData)

```

## Effective Confidence interval `r low` to `r high`

# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 
```{r}
MyData2 = numeric(1000)
for (i in 1:1000) {
  MyData2[i]= BrLw(odds = 4.36)
}
sMyData2 = sort(MyData2) 
low=sMyData2[25]
high=sMyData2[975]
average=mean(sMyData2)
```

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->
To find the fair odds for Professor Brownlow’s bets, we ran 1000 simulations of his earnings using different odds values. For each set of simulations, we calculated the average earnings. We adjusted the odds until the average earnings were close to zero, meaning he would break even in the long run. Once we found the odds that gave an average around zero, we sorted the results and took the 25th and 975th values to create a 95% confidence interval. This shows the range where most of his earnings would fall, even when betting at fair odds.



# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.
```{r}
MyData3= numeric(1000)
for (i in 1:1000) {
  MyData2[i]=BrLw(numClass = 28)
}
sMyData3 =sort(MyData3)
low=sMyData3[25]
high=sMyData3[975]
middle=mean(sMyData3)[500]

```


<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

We ran 1000 simulations of Professor Brownlow’s betting outcomes for different class sizes using the BrLw() function. For each class size, we calculated the average earnings from the 1000 simulations. We manually adjusted the number of students in the class until the average earnings came close to zero, which would mean the bet is fair at that class size.

After finding the fair class size, we sorted the earnings and calculated the 2.5th and 97.5th percentiles (values 25 and 975 from the sorted list) to form a 95% confidence interval. This gives a range of how much Professor Brownlow might win or lose, even when betting at fair class size.


# How much money would Brownlow be expected to end up with if the bets went as described.
He made in a 95% confidence interval with an interval from $226-337 dollars with an average of 281.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->
If Professor Brownlow is consistently making or losing money, he can make the bet fair by adjusting either the payout or the class size. To balance his profits around zero, he should change the payout ratio. If he's making too much money, he should increase the payout to students. If he's losing money, he should decrease the payout. Alternatively, he can keep the 2:1 payout but adjust the class size. Through simulation, we found that when the class size is around 28 students, the bet becomes fair and his expected earnings center around zero. By tweaking either the payout or the number of students, Professor Brownlow can create a fair betting system.

